\textbf{SPECFEM3D}\newline \textbf{Cartesian}\newline \textbf{User Manual}

### 2022-11-12 doc/USER_MANUAL/07_adjoint_simulations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

After the adjoint simulation has completed successfully, collect the
seismograms from \texttt{LOCAL\_PATH}.
\begin{itemize}
\item These adjoint seismograms are recorded at the locations of the original
earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
have names of the form \texttt{NT.S?????.S??.sem} for the six-component
strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
and ~\\
 \texttt{NT.S?????.BX?.sem} for the three-component displacements
(\texttt{BXN,BXE,BXZ}) recorded at these locations.
\item \texttt{S?????} denotes the source number; for example, if the original
\texttt{CMTSOLUTION} provides only a point source, then the seismograms
collected will start with \texttt{S00001}.
\item These adjoint seismograms provide critical information for the computation
of the gradient of the misfit function.
\end{itemize}
\end{enumerate}

>>>>

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\newline
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

After the adjoint simulation has completed successfully, collect the
seismograms from \texttt{LOCAL\_PATH}.
\begin{itemize}
\item These adjoint seismograms are recorded at the locations of the original
earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
have names of the form \texttt{NT.S?????.S??.sem} for the six-component
strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
and ~\newline
 \texttt{NT.S?????.BX?.sem} for the three-component displacements
(\texttt{BXN,BXE,BXZ}) recorded at these locations.
\item \texttt{S?????} denotes the source number; for example, if the original
\texttt{CMTSOLUTION} provides only a point source, then the seismograms
collected will start with \texttt{S00001}.
\item These adjoint seismograms provide critical information for the computation
of the gradient of the misfit function.
\end{itemize}
\end{enumerate}

### 2022-11-12 doc/USER_MANUAL/08_doing_tomography.tex
switched the license to GNU version 3 (now that we use pyCMT3D, which is LGPL v3, the combination automatically becomes GNU version 3). Also added FWI inversion information to the users manual.
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@

Besides the ability to compute adjoint sensitivity kernels, the SPECFEM3D package provides a number of tools useful for post-processing kernels, gradients and conducting tomographic model updates. The tomographic and post-processing tools can be compiled by typing in the main directory:
\begin{verbatim}
make tomography
make postprocess
\end{verbatim}
You can modify specific settings affecting, e.g., the (isotropic) parameterization or directory setup structure in file \texttt{setup/constants\_tomography.h}. The compiled binaries can be found in directory \texttt{bin/}.\\

@@

Please note that the tomographic routines are very similar for both SPECFEM3D\_Cartesian and SPECFEM3D\_GLOBE versions. Some differences occur, e.g., when reading in mesh files and when the maximum of the gradient is taken for the update step length. Also note that currently only the binary database format is supported, no ADIOS file format support is implemented yet. In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\\

@@

\item [-]{\bf Output:}
The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\\
\end{itemize}

@@

\noindent
{\bf Kernel names}: The summation tool will look for kernels with following names:
\begin{itemize}
\item [-] for an {\it isotropic} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_alpha_kernel.bin
proc***_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_alpha_kernel.bin
proc***_reg1_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_kernel.bin
proc***_bulk_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_kernel.bin
proc***_reg1_bulk_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_c_kernel.bin
proc***_bulk_betav_kernel.bin
proc***_bulk_betah_kernel.bin
proc***_eta_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_c_kernel.bin
proc***_reg1_bulk_betav_kernel.bin
proc***_reg1_bulk_betah_kernel.bin
proc***_reg1_eta_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

Note that these event kernels are stored after a kernel simulation (\texttt{SIMULATION\_TYPE = 3}) in the \texttt{LOCAL\_PATH} directory, which by default points to directory \texttt{DATABASES\_MPI/}. Isotropic kernels will be created by default. To create transversely isotropic kernels, you set \texttt{ANISOTROPIC\_KL} and \texttt{SAVE\_TRANSVERSE\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}. \\

\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_hess_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_hess_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

These Hessian kernels approximate the diagonal elements of the Hessian matrix and can be used as preconditioner in a gradient optimization scheme.
To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\\

@@

For SPECFEM3D\_GLOBE, by default only kernels in the crust/mantle region ("reg1") will be considered for summation. You can change the default by setting parameter \texttt{REG} in file \texttt{constants\_tomography.h} to the region of interest. \\

@@

Note that instead of using the density kernels $K_{\rho}$ for model updates, setting \texttt{USE\_RHO\_SCALING} to \texttt{.true.} will
ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\\

>>>>

\red{\Large ANNOUNCEMENT (June 2017): See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline

@@

Besides the ability to compute adjoint sensitivity kernels, the SPECFEM3D package provides a number of tools useful for post-processing kernels, gradients and conducting tomographic model updates. The tomographic and post-processing tools can be compiled by typing in the main directory:
\begin{verbatim}
make tomography
make postprocess
\end{verbatim}
You can modify specific settings affecting, e.g., the (isotropic) parameterization or directory setup structure in file \texttt{setup/constants\_tomography.h}. The compiled binaries can be found in directory \texttt{bin/}.\newline

@@

Please note that the tomographic routines are very similar for both SPECFEM3D\_Cartesian and SPECFEM3D\_GLOBE versions. Some differences occur, e.g., when reading in mesh files and when the maximum of the gradient is taken for the update step length. Also note that currently only the binary database format is supported, no ADIOS file format support is implemented yet. In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\newline

@@

\item [-]{\bf Output:}
The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\newline
\end{itemize}

@@

\noindent
{\bf Kernel names}: The summation tool will look for kernels with following names:
\begin{itemize}
\item [-] for an {\it isotropic} model parameterization, kernel names are:\newline
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_alpha_kernel.bin
proc***_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_alpha_kernel.bin
proc***_reg1_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\newline

\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\newline
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_kernel.bin
proc***_bulk_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_kernel.bin
proc***_reg1_bulk_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\newline

\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\newline
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_c_kernel.bin
proc***_bulk_betav_kernel.bin
proc***_bulk_betah_kernel.bin
proc***_eta_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_c_kernel.bin
proc***_reg1_bulk_betav_kernel.bin
proc***_reg1_bulk_betah_kernel.bin
proc***_reg1_eta_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\newline

Note that these event kernels are stored after a kernel simulation (\texttt{SIMULATION\_TYPE = 3}) in the \texttt{LOCAL\_PATH} directory, which by default points to directory \texttt{DATABASES\_MPI/}. Isotropic kernels will be created by default. To create transversely isotropic kernels, you set \texttt{ANISOTROPIC\_KL} and \texttt{SAVE\_TRANSVERSE\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}. \newline

\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\newline
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_hess_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_hess_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\newline

These Hessian kernels approximate the diagonal elements of the Hessian matrix and can be used as preconditioner in a gradient optimization scheme.
To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\newline

@@

For SPECFEM3D\_GLOBE, by default only kernels in the crust/mantle region ("reg1") will be considered for summation. You can change the default by setting parameter \texttt{REG} in file \texttt{constants\_tomography.h} to the region of interest. \newline

@@

Note that instead of using the density kernels $K_{\rho}$ for model updates, setting \texttt{USE\_RHO\_SCALING} to \texttt{.true.} will
ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\newline

### 2022-11-12 doc/USER_MANUAL/09_noise_simulations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

As usual, the three main input files are crucial: \texttt{Par\_file},
\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
those input files should be located in directory \texttt{DATA/}.\\

\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
it may seem unexpected to have it here, since the noise simulations
should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
However, for noise simulations, it is critical to have no earthquakes.
In other words, the moment tensor specified in \texttt{CMTSOLUTION}
must be set to zero manually!\\

\texttt{STATIONS} remains the same as in previous earthquake simulations,
except that the order of receivers listed in \texttt{STATIONS} is
now important. The order will be used to determine the `main' receiver,
i.e., the one that simultaneously cross correlates with the others.\\

\texttt{Par\_file} also requires careful attention. A parameter called
\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
equals 0, a regular earthquake simulation will be run. When it is
1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
Should you be confused by the three steps, refer to \citet{trompetal2010}
for details.\\

Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
While for regular earthquake simulations this parameter specifies
the length of synthetic seismograms generated, for noise simulations
it specifies the length of the seismograms used to compute cross correlations.
The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
is not zero.\\

There are other parameters in \texttt{Par\_file} which should be given
specific values. For instance, since the first two steps for calculating
noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
have to reconstruct the ensemble forward wavefields in adjoint simulations,
therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\

@@

It is better to run the three steps continuously within
the same job on a cluster, otherwise you have to collect the saved surface movies
from the old nodes to the new nodes. This process varies
from cluster to cluster and thus cannot be discussed here. Please
ask your cluster administrator for information/configuration of the
cluster you are using.\\

@@

After those simulations, you have all stuff you need, either in the
\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
in \texttt{Par\_file} (which are most probably on local nodes).
Collect whatever you want from the local nodes to your workstation,
and then visualize them. This process is the same as what you may
have done for regular earthquake simulations. Refer to other chapters
if you have problems.\\

Simply speaking, two outputs are the most interesting: the simulated
ensemble cross correlations and one contribution of the ensemble sensitivity
kernels.\\

The simulated ensemble cross correlations are obtained after the second
simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
the simulated ensemble cross correlations. Collect them immediately
after Step 2, or the Step 3 will overwrite them. Note that we have
a `main' receiver specified by \texttt{irec\_main\_noise}, the
seismogram at one station corresponds to the cross correlation between
that station and the `main'. Since the seismograms have three components,
we may obtain cross correlations for different components as well,
not necessarily the cross correlations between vertical components.\\

One contribution of the ensemble cross-correlation sensitivity kernels
are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
on local nodes. The ensemble kernel files are named the same as regular
earthquake kernels.\\

@@

In order to illustrate noise simulations in an easy way, one example
is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
for explanations. \\

Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
which has at least 4 cores with `mpif90' working properly. \\

If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
using:\\

\texttt{./pre-processing.sh}\\

>>>>

As usual, the three main input files are crucial: \texttt{Par\_file},
\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
those input files should be located in directory \texttt{DATA/}.\newline

\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
it may seem unexpected to have it here, since the noise simulations
should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
However, for noise simulations, it is critical to have no earthquakes.
In other words, the moment tensor specified in \texttt{CMTSOLUTION}
must be set to zero manually!\newline

\texttt{STATIONS} remains the same as in previous earthquake simulations,
except that the order of receivers listed in \texttt{STATIONS} is
now important. The order will be used to determine the `main' receiver,
i.e., the one that simultaneously cross correlates with the others.\newline

\texttt{Par\_file} also requires careful attention. A parameter called
\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
equals 0, a regular earthquake simulation will be run. When it is
1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
Should you be confused by the three steps, refer to \citet{trompetal2010}
for details.\newline

Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
While for regular earthquake simulations this parameter specifies
the length of synthetic seismograms generated, for noise simulations
it specifies the length of the seismograms used to compute cross correlations.
The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
is not zero.\newline

There are other parameters in \texttt{Par\_file} which should be given
specific values. For instance, since the first two steps for calculating
noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
have to reconstruct the ensemble forward wavefields in adjoint simulations,
therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\newline

@@

It is better to run the three steps continuously within
the same job on a cluster, otherwise you have to collect the saved surface movies
from the old nodes to the new nodes. This process varies
from cluster to cluster and thus cannot be discussed here. Please
ask your cluster administrator for information/configuration of the
cluster you are using.\newline

@@

After those simulations, you have all stuff you need, either in the
\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
in \texttt{Par\_file} (which are most probably on local nodes).
Collect whatever you want from the local nodes to your workstation,
and then visualize them. This process is the same as what you may
have done for regular earthquake simulations. Refer to other chapters
if you have problems.\newline

Simply speaking, two outputs are the most interesting: the simulated
ensemble cross correlations and one contribution of the ensemble sensitivity
kernels.\newline

The simulated ensemble cross correlations are obtained after the second
simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
the simulated ensemble cross correlations. Collect them immediately
after Step 2, or the Step 3 will overwrite them. Note that we have
a `main' receiver specified by \texttt{irec\_main\_noise}, the
seismogram at one station corresponds to the cross correlation between
that station and the `main'. Since the seismograms have three components,
we may obtain cross correlations for different components as well,
not necessarily the cross correlations between vertical components.\newline

One contribution of the ensemble cross-correlation sensitivity kernels
are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
on local nodes. The ensemble kernel files are named the same as regular
earthquake kernels.\newline

@@

In order to illustrate noise simulations in an easy way, one example
is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
for explanations. \newline

Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
which has at least 4 cores with `mpif90' working properly. \newline

If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
using:\newline

\texttt{./pre-processing.sh}\newline

### 2022-11-12 doc/USER_MANUAL/10_gravity_calculations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
That feature is still experimental but should work just fine.\\

To see how it is implemented and to use it, type this in the root directory of the code:\\

\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}

>>>>

SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
That feature is still experimental but should work just fine.\newline

To see how it is implemented and to use it, type this in the root directory of the code:
\begin{verbatim}
grep -i GRAVITY_INTEGRALS src/*/*
\end{verbatim}

### 2022-11-12 doc/USER_MANUAL/13_changing_the_model.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\\
 \indent (1) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= tomo}, \\
 or for more user control: \\
 \indent (2) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{verbatim}
domain_ID material_ID tomography elastic file_name 1
\end{verbatim}
where: \\
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\\
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \\
 \indent \texttt{tomography} keyword for tomographic material definition,
\\
 \indent \texttt{elastic} keyword for elastic material definition,
\\
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\\

>>>>

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\newline
 \indent (1) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= tomo}, \newline
 or for more user control: \newline
 \indent (2) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{verbatim}
domain_ID material_ID tomography elastic file_name 1
\end{verbatim}
where: \newline
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\newline
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \newline
 \indent \texttt{tomography} keyword for tomographic material definition,
\newline
 \indent \texttt{elastic} keyword for elastic material definition,
\newline
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\newline

### 2022-11-12 doc/USER_MANUAL/14_post_processing.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

In order to convert between SAC format and ASCII files, useful scripts
are provided in the subdirectories ~\\
 \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.

@@

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation ~\\
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
code to convert the binary shakemap data into GMT xyz files. A sample
script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
and then plot the shakemaps in GMT, for example:
\begin{verbatim}
plot_shakemap.gmt.pl data _dir type(1,2,3) CMTSOLUTION
\end{verbatim}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

>>>>

In order to convert between SAC format and ASCII files, useful scripts
are provided in the subdirectories ~\newline
 \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.

@@

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation ~\newline
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
code to convert the binary shakemap data into GMT xyz files. A sample
script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
and then plot the shakemaps in GMT, for example:
\begin{verbatim}
plot_shakemap.gmt.pl data _dir type(1,2,3) CMTSOLUTION
\end{verbatim}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

### 2022-11-12 doc/USER_MANUAL/authors.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

\section*{https://specfem3d.readthedocs.io/en/latest/authors/}

@@

%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
Nikhil Vaidyar

The cover graphic of the manual was created by Santiago Lombeyda from
Caltech's Center for Advanced Computing Research (CACR), USA, with free satellite clipart pictures
from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\\

The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\\

>>>>

\section*{Authors}

@@

%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
Since then it has been developed and maintained by a development team: in alphabetical order,
Michael Afanasiev,
Jean-Paul (Pablo) Ampuero,
\'Etienne Bachmann,
Kangchen Bai,
Piero Basini,
C\'eline Blitz,
Alexis Bottero,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Paul Cristini,
Cl\'ement Durochat,
Percy Galvez,
Hom Nath Gharti,
Dominik G\"oddeke,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Matthieu Lefebvre,
Qinya Liu,
Youshan Liu,
David Luet,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Vadim Monteiller,
Surendra Nadh Somala,
Masaru Nagaso,
Tarje Nissen-Meyer,
Daniel Peter,
Kevin Pouget,
Max Rietmann,
Elliott Sales de Andrade,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
James Smith,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Brice Videau,
Jean-Pierre Vilotte,
Zhinan Xie,
Chang-Hua Zhang,
Hejun Zhu.\newline

The cover graphic of the manual was created by Santiago Lombeyda from
Caltech's Center for Advanced Computing Research (CACR), USA, with free satellite clipart pictures
from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\newline

The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\newline

### 2022-11-12 doc/USER_MANUAL/notes_and_acknowledgement.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
is an open-source visualization platform.

Please e-mail your feedback, questions, comments, and suggestions
to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

>>>>

OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
is an open-source visualization platform.\newline

Please provide your feedback, questions, comments, and suggestions through the github repository.
